---
title: "Assignment 5"
author: "Niteshyadav sanna"
date: "2024-04-07"
output:
  html_document:
    df_print: paged
---
##Load Data Set and Libraries
```{r}
library(cluster)
library(ISLR)
library(caret)
library(dplyr)
library(tidyverse)
library(factoextra)
library(ggplot2)
library(proxy)
library(NbClust)
library(ppclust)
library(dendextend)
library(tinytex)
```

```{r}
cereal <- read.csv("C:/Users/sanna/Downloads/Cereals.csv")
```
##Review Data Structure
```{r}
head(cereal)
```
```{r}
str(cereal)
```
```{r}
summary(cereal)
```
##Data Preprocessing
```{r}
cereal_scaled <- cereal
```
```{r}
cereal_scaled[ , c(4:16)] <- scale(cereal[ , c(4:16)])
```
```{r}
cereal_preprocessed <- na.omit(cereal_scaled)
```
```{r}
head(cereal_preprocessed)
```
##Assignment Task A

```{r}
cereal_d_euclidean <- dist(cereal_preprocessed[ , c(4:16)], method =
"euclidean")
```
```{r}
ag_hc_single <- agnes(cereal_d_euclidean, method = "single")
```
```{r}
plot(ag_hc_single,
 main = "Customer Cereal Ratings - AGNES - Single Linkage Method",
 xlab = "Cereal",
 ylab = "Height",
 cex.axis = 1,
 cex = 0.55,
 hang = -1)
```
```{r}
ag_hc_complete <- agnes(cereal_d_euclidean, method = "complete")
```
```{r}
plot(ag_hc_complete,
 main = "Customer Cereal Ratings - AGNES - Complete Linkage Method",
 xlab = "Cereal",
 ylab = "Height",
 cex.axis = 1,
 cex = 0.55,
 hang = -1)

```
```{r}
ag_hc_average <- agnes(cereal_d_euclidean, method = "average")
```
```{r}
plot(ag_hc_average,
 main = "Customer Cereal Ratings - AGNES - Average Linkage Method",
 xlab = "Cereal",
 ylab = "Height",
 cex.axis = 1,
 cex = 0.55,
 hang = -1)

```
```{r}
ag_hc_ward <- agnes(cereal_d_euclidean, method = "ward")
```
```{r}
plot(ag_hc_ward,
 main = "Customer Cereal Ratings - AGNES - Ward Linkage Method",
 xlab = "Cereal",
 ylab = "Height",
 cex.axis = 1,
 cex = 0.55,
 hang = -1)
```
##Assignment Task B

```{r}
fviz_nbclust(cereal_preprocessed[ , c(4:16)], hcut, method = "wss", k.max =
25) +
 labs(title = "Optimal Number of Clusters - Elbow Method") +
 geom_vline(xintercept = 12, linetype = 2)
```
```{r}
fviz_nbclust(cereal_preprocessed[ , c(4:16)],
 hcut,
method = "silhouette",
 k.max = 25) +
 labs(title = "Optimal Number of Clusters - Silhouette Method")
```
```{r}
plot(ag_hc_ward,
 main = "AGNES - Ward Linkage Method - 12 Clusters Outlined",
 xlab = "Cereal",
 ylab = "Height",
 cex.axis = 1,
 cex = 0.55,
 hang = -1)
rect.hclust(ag_hc_ward, k = 12, border = 1:12)
```
##Assignment Task C

```{r}
ward_clusters_12 <- cutree(ag_hc_ward, k = 12)
```
```{r}
cereal_preprocessed_1 <- cbind(cluster = ward_clusters_12,
cereal_preprocessed)
```
```{r}
set.seed(982579)
```
```{r}
cerealIndex <- createDataPartition(cereal_preprocessed$protein, p=0.3, list =
F)
cereal_preprocessed_PartitionB <- cereal_preprocessed[cerealIndex, ]
cereal_preprocessed_PartitionA <- cereal_preprocessed[-cerealIndex,] 
```
```{r}
cereal_d_euclidean_A <- dist(cereal_preprocessed_PartitionA[ , c(4:16)],
method = "euclidean")
```
```{r}
ag_hc_ward_A <- agnes(cereal_d_euclidean_A, method = "ward")
```
```{r}
plot(ag_hc_ward_A,
 main = "Customer Cereal Ratings - Ward Linkage Method - Partition A",
 xlab = "Cereal",
 ylab = "Height",
 cex.axis = 1,
 cex = 0.55,
 hang = -1)
```
```{r}
ward_clusters_12_A <- cutree(ag_hc_ward_A, k = 12)
```
```{r}
cereal_preprocessed_A <- cbind(cluster = ward_clusters_12_A,
cereal_preprocessed_PartitionA)
```
```{r}
ward_Centroids_A <- aggregate(cereal_preprocessed_A[ , 5:17],
list(cereal_preprocessed_A$cluster), mean)
ward_Centroids_A <- data.frame(Cluster = ward_Centroids_A[ , 1], Centroid =
rowMeans(ward_Centroids_A[ , -c(1:4)]))
ward_Centroids_A <- ward_Centroids_A$Centroid
```
```{r}
cereal_preprocessed_PartitionB_centers <-
data.frame(cereal_preprocessed_PartitionB[, 1:3], Center =
rowMeans(cereal_preprocessed_PartitionB[ , 4:16]))
```
```{r}
B_to_A_centers <- dist(ward_Centroids_A,
cereal_preprocessed_PartitionB_centers$Center, method = "euclidean")
```
```{r}
cereal_preprocessed_B <- cbind(cluster =
c(4,8,7,3,5,6,7,11,11,10,8,5,10,1,10,1,4,12,12,7,7,1,4,9),
cereal_preprocessed_PartitionB)
```
```{r}
cereal_preprocessed_2 <- rbind(cereal_preprocessed_A, cereal_preprocessed_B)
cereal_preprocessed_1 <-
cereal_preprocessed_1[order(cereal_preprocessed_1$name), ]
cereal_preprocessed_2 <-
cereal_preprocessed_2[order(cereal_preprocessed_2$name), ]
```
```{r}
sum(cereal_preprocessed_1$cluster == cereal_preprocessed_2$cluster)
```
```{r}
ggplot(data = cereal_preprocessed_1, aes(cereal_preprocessed_1$cluster)) +
 geom_bar(fill = "yellow3") +
 labs(title="Count of Cluster Assignments - All Original Data") +
 labs(x="Cluster Assignment", y="Count") +
 guides(fill=FALSE) +
 scale_x_continuous(breaks=c(1:12)) +
 scale_y_continuous(breaks=c(5,10,15,20), limits = c(0,25))
```
```{r}
ggplot(data = cereal_preprocessed_2, aes(cereal_preprocessed_2$cluster)) +
 geom_bar(fill = "yellow3") +
 labs(title="Count of Cluster Assignments - Partitioned Data") +
 labs(x="Cluster Assignment", y="Count") +
 guides(fill=FALSE) +
 scale_x_continuous(breaks=c(1:12)) +
 scale_y_continuous(breaks=c(5,10,15,20), limits = c(0,25))
```
#Using the partitioned data, we can see graphically that Cluster 3 shrank significantly.Some of the other clusters became larger as a result. The graph indicates that the clusters split when the data was dispersed more equally among the 12 clusters.

##Assignment Task D

#“The public elementary schools would like to select a selection of cereals to be served in their cafeterias on a daily basis. There is a different cereal available each day, but all of the cereals should encourage a balanced diet. You are to locate a cluster of "healthy cereals" in order to complete this mission. Is it necessary to normalize the data? If not, how ought they to be applied in the cluster analysis?

#In this instance, normalizing the data would not be suitable. This is because the way the nutrition data from cereals is scaled or normalized depends on the specific cereal sample that is being studied. This means that only cereals with exceptionally high sugar content and extremely low iron, fiber, and other nutritional content may have been included in the data set that was collected. The nutritional value of cereal for a child cannot be ascertained when the data is standardized or scaled inside the sample set. Uninformed observers may conclude that cereal with an iron score of 0.999 almost completely satisfies a child's iron requirements; yet, it may simply be the best cereal in the sample set with negligible to no nutritious value.

#Thus, a more acceptable way to handle the data would be to preprocess it as a ratio to a child's daily recommended consumption of calories, fiber, carbohydrates, etc. Consequently, a few key variables would not be able to taint the distance estimates and analysts would be better equipped to examine the clusters. To determine how much XX cereal a student would need to eat each day to achieve prescribed nutritional needs, an analyst analyzing the clusters may look at the cluster average. This would give the staff members the ability to select the "healthy" cereal clusters from an informed menu.













